Storm surges in the Strait of Georgia simulated with a regional model

### 2015-05-05 AO_guides/manuscriptAO.tex
removed commetn about northern forcing

In a simulation without atmospheric forcing, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side (Figure \ref{fig:spatial_sshonly}). While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude. Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.  Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps because the Strait of Georgia is also affected by the surge entering the domain from Johnstone Strait at the northern boundary or the incoming surge wave is partially reflected from the mainland coast.

@@

Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated. In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases and generally higher skill scores for the water level heights.  Further, in a single simulation, the skill scores are fairly consistent between the four stations examined. We have also assessed the relative importance of different forcing factors contributing to the storm surge. The sea surface height anomaly entering the system from the Pacific Ocean, or remote forcing,  contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}. Although the large scale wind patterns and atmospheric conditions are important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude. However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season. Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due to anomalies entering from both the northern and southern regions or a partial reflection from the mainland coast. Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands. The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.

>>>>

In a simulation without atmospheric forcing, the surge amplitude across the Strait of Georgia is uniform with no pronounced difference between the mainland side and Vancouver Island side (Figure \ref{fig:spatial_sshonly}). While the surge entering the domain from the Pacific Ocean is the leading order contribution, the winds may induce localized changes to the surge amplitude. Interestingly, the winds over the Strait of Juan de Fuca cause a decrease in surge amplitude in this region as the winds are acting to push the water towards the mainland.  {\color{red} Even without atmospheric forcing, there is still a significant difference in surge amplitude between the Strait of Georgia and the Strait of Juan de Fuca, perhaps the incoming surge wave is partially reflected from the mainland coast in the Strait of Georgia}.

@@

Hindcasts of five significant events in February 2006, November 2006, December 2006, November 2009, and December 2012 were simulated. In general, the model performs well at reproducing the observed water level heights and residuals, with Willmott skill scores greater than 0.88 in all cases and generally higher skill scores for the water level heights.  Further, in a single simulation, the skill scores are fairly consistent between the four stations examined. We have also assessed the relative importance of different forcing factors contributing to the storm surge. The sea surface height anomaly entering the system from the Pacific Ocean, or remote forcing,  contributes most significantly to the surge amplitude, in agreement with previous studies in this region \citep{murty1995storm}. Although the large scale wind patterns and atmospheric conditions are important in setting up the surge on the Pacific Ocean boundaries, the local winds within the Strait of Georgia are not the leading order contribution to surge amplitude. However, the local wind patterns can set up small spatial differences in the surge, resulting in higher surges on the mainland side of the Strait of Georgia due to the usually southerly winds during storm surge season. {\color{red}Additionally, surge amplitudes in the Strait of Georgia are typically higher than those in the Strait of Juan de Fuca, perhaps due a partial reflection of the surge from the mainland coast.} Results from an additional simulation with no tidal forcing but only surge propagation and atmospheric forcing found that nonlinear effects due to the tide-surge interaction were most important north of the San Juan and Gulf Islands. The modelled surge amplitude increased by about 10\% at Point Atkinson when the tide-surge interaction was neglected in one example.
